1. A brewery-filling machine is adjusted to fill quart bottles
with a mean of 32.00 oz. of ale and a variance of 0.0016.
Assume the amount of fill is normally distributed.
$$\mu=32$$
$$\sigma^2=0.0016$$

\bigskip
{\bf (a) How many percent of the bottles contains more than 32.02 oz.?}

\bigskip
We need to find the probability that a bottle of beer has more than 32.02 ounces.
$$P(X>32.02)=1-P(X\le32.02)$$
To do the calculation, we have to switch to the $Z$ domain.
$$z={x-\mu\over\sigma}={x-32\over0.04}$$
Hence
$$1-P(X\le32.02)=1-P(Z\le0.5)=0.3085$$
Check using Excel.
$$1-\hbox{NORMDIST(32.02,32,0.04,TRUE)}=0.3085$$
Multiply by 100 to get the percentage 31\%.

\bigskip
{\bf (b) If 10 quart bottles are randomly selected, what is the
probability that more than two quart bottles contain more
than 32.04 oz. of ale?}

\bigskip
First, find the probability that 1 bottle contains
more than 32.04 oz. of ale.
$$1-\hbox{NORMDIST(32.04,32,0.04,TRUE)}=0.158655$$
Next we have to work out a discrete distribution.
What kind of distribution is it?
A binomial distribution?
We can check by typing BINOMDIST into Excel and looking
at the arguments.
We see that the arguments are ``number\_s, trials, probability\_s, cumulative''
where $s$ means success.
Ok, let's give BINOMDIST a try.
$$P(X>2)=1-P(X\le2)=1-\hbox{BINOMDIST(2,10,0.158655,TRUE)}=0.202753$$

\vfill
\eject

{\bf (c) If quart bottles are randomly selected, what is the
probability that at most five quart bottles are needed to have
the first quart bottle contain less than 31.98 oz. of ale?}

\bigskip
Again, start by working out the probability for 1 bottle.
$$P(X\le31.98)=\hbox{NORMDIST(31.98,32,0.04,TRUE)}=0.308538$$
Here is a summary of the Excel functions...
$$\matrix{
\hbox{BINOMDIST} & \hbox{number\_s} & \hbox{trials} & \hbox{probability\_s} & \hbox{cumulative}\cr\cr
\hbox{HYPGEOMDIST} & \hbox{sample\_s} & \hbox{number\_sample} & \hbox{population\_s}
& \hbox{number\_pop}\cr\cr
\hbox{NEGBINOMDIST}&
\hbox{number\_f}&
\hbox{number\_s}&
\hbox{probability\_s}\cr\cr
\hbox{POISSON}&
\hbox{x}&
\hbox{mean}&
\hbox{cumulative}
}$$
We see that we can work out the answer by summing over number\_f 
(number of failures) in NEGBINOMDIST.
$$P(X\le5)
=\hbox{NEGBINOMDIST(0,1,0.3085)}+\cdots
+\hbox{NEGBINOMDIST(4,1,0.3085)}=0.8419
$$

\bigskip
{\bf (d) If quart bottles are randomly selected, what is the probability
that at most five quart bottles are needed to have the second quart bottle
contain more than 31.98 oz. and less than 32.04 oz. of ale?}

$$P(31.98<X<32.04)=\hbox{NORMDIST(32.04,32,0.04,TRUE)}
-\hbox{NORMDIST(31.98,32,0.04,TRUE)}=0.5328$$
Again, sum over number\_f but this time with number\_s set to 2.
$$P(X\le5)=
\hbox{NEGBINOMDIST(0,2,0.5328)}+\cdots
+\hbox{NEGBINOMDIST(3,2,0.5328)}=0.8508
$$

\end