\beginsection 1.

The Webster National Bank is reviewing its service charge and interest
payment policies on checking accounts.
The bank has found the mean and standard deviation of the daily balance
on personal checking accounts is \$550.00 and \$150.00, respectively.
In addition, the daily balances are normally distributed.
$$\mu=550$$
$$\sigma=150$$
$$z={x-550\over150}$$

\beginsection 1a.

Percent of customers with daily balance
below \$790.00?

\bigskip
Solution:
$$P(Z<790)=100\times\hbox{\tt NORMDIST(790,550,150,TRUE)}=94.52\%$$

\beginsection 1b.

Percent of customers with balances between \$220 and \$784?

\bigskip
Solution:
$$P(220<Z<784)=100\times(\hbox{\tt NORMDIST(784,550,150,TRUE)}
-\hbox{\tt NORMDIST(220,550,150,TRUE})=92.67\%$$

\beginsection 1c.

Percent of customer's with balances in excess of \$211?

\bigskip
Solution:
$$P(Z>211)=1-P(Z<211)=1-\hbox{\tt NORMDIST(211,550,150,TRUE)}=98.81\%$$

\beginsection 1d.

What balance is exceeded by 10.75\% of the customers?

\bigskip
Solution:
$$P(Z>a)=0.1075,\qquad P(Z<a)=1-0.1075=0.8925,\qquad a=\hbox{\tt NORMINV(0.8925,550,150)}
=\$735.99$$

\beginsection 1e.

What balance is not exceeded by 20.05\% of the customers?

\bigskip
Solution:
$$P(Z<a)=0.2005\,\qquad a=\hbox{\tt NORMINV(0.2005,550,150)}=\$424.02$$

\vfill
\eject

\beginsection 2.

Assume that the width of the steel beams produced by a machine is normally
distributed with mean 10 inches and standard deviation 1.5 inches.
$$\mu=10$$
$$\sigma=1.5$$
$$z={x-10\over1.5}$$

\beginsection 2a.

What is the probability that at least 10 of 12 randomly selected steel
beams have width between 8.71 and 13.855 inches?

\bigskip
Solution:
$$P(8.71<Z<13.855)=\hbox{\tt NORMDIST(13.855,10,1.5,TRUE)}-
\hbox{\tt NORMDIST(8.71,10,1.5,TRUE)}=0.8000$$
Next, use the binomial distribution.
$$P(X\le10)=\hbox{\tt BINOMDIST(10,12,0.8,TRUE)}=0.7251$$

\beginsection 2b.

What is the probability that the total width of four randomly selected steel
beams is more than 44.5 inches?

\bigskip
Solution:
$$\mu=(4)(10)=40\,\qquad \sigma=(\sqrt4)(1.5)=3$$
$$P(Y>44.5)=1-P(Y<44.5)=1-\hbox{\tt NORMDIST(44.5,40,3,TRUE)}=0.0668$$

\beginsection 2c.

How many steel beams do we expect to select until we find one that has a width
of more than 13 inches?

\bigskip
Solution:
$$P(Z>13)=1-P(Z<13)=1-\hbox{\tt NORMDIST(13,10,1.5,TRUE)}=0.0228$$
Next, use the mean of a geometric distribution.
$$E(X)={1-0.0228\over0.0228}=42.86$$

\vfill
\eject

\beginsection 3.

The time to failure of a personal computer display tube is exponentially
distributed with mean 3 years.
$$\theta=3$$

\beginsection 3a.

A company offers insurance on these tubes for the first year of usage.
On what percentage of policies will they have to pay a claim?

\bigskip
Solution:
$$100\%\times P(X<1)=100\%\times(1-e^{-t/\theta})=100\%\times(1-e^{-1/3})=28.35\%$$

\beginsection 3b.

If a user orders three tubes (two as spares) what is the probability
that all will fail within six years?

\bigskip
Solution: The average unit of measurement for 3 consecutive failures is 9 years.
$$\theta=9$$
$$P(X<6)=1-e^{-6/9}=0.4866$$
Check with Excel.
$$\hbox{\tt EXPONDIST(6,1/9,TRUE)}=0.4866$$

\beginsection 4.

Let $U$ be a uniformly distributed random variable with mean 2 and variance 4/3,
what is the probability that the roots of $Y^2+4UY+(U+1)=0$ are real?

\bigskip
Solution: Since $U$ is uniformly distributed we have to find $a$ and $b$.
$$\mu={b+a\over2}=2,\qquad a=4-b$$
$$\sigma^2={(b-a)^2\over12}={4\over3},\qquad b^2-2ab+a^2-16=0$$
Hence
$$a=0,\qquad b=4$$
The distribution function is
$$F(t)=\cases{
0, & $t<0$\cr
\cr
t/4, & $0\le t<4$\cr
\cr
1, & $4\le t$\cr
}$$
The roots of the equation are
$$-2U\pm{1\over2}\sqrt{16U^2-4U-4}$$
The roots of the equation are real for
$$U<-0.3904,\qquad U>0.6404$$
$$P(X>0.6404)=1-P(X<0.6404)=1-{0.6404\over4}=0.8399$$

\vfill
\eject

\beginsection 5.

The replenishment time for a certain product is known to be gamma
distributed with mean 40 and variance 400.
Find the probability that an order is received within the first
20 days after it is ordered.

\bigskip
Solution: We are given $\mu$ and $\sigma^2$.
$$\mu=\alpha\theta=40,\qquad\sigma^2=\alpha\theta^2=400$$
Hence
$$\alpha=4,\qquad\theta=10$$

$$P(X<20)=\hbox{\tt GAMMADIST(20,1,10,TRUE)}=0.8647$$
$$P(X<20)=\hbox{\tt EXPONDIST(20,1/10,TRUE)}=0.8647$$

\beginsection 6.

Let random variable $X$ be uniformly distributed on the interval $(0,1)$.
Find the probability density function of $Y$, if $Y=X^{1/\alpha}$,
where $\alpha>0$.

\bigskip
Solution: We have
$$Y=F^{-1}(X)=X^{1/\alpha}$$
Solve for $F$.
$$y=x^{1/\alpha},\qquad y^\alpha=x$$
Hence
$$X=F(Y)=Y^\alpha$$
Now we can take the derivative of $F$ to get the probability density function for $Y$.
$$f(y)={dF\over dy}=\alpha y^{\alpha-1}$$

\vfill
\eject

\beginsection 7.

Consider the random variable $X$ with probability density function
$$f(x)=\cases{
\displaystyle{
{\Gamma(\alpha+\beta)\over\Gamma(\alpha)\Gamma(\beta)}x^{\alpha-1}(1-x)^{\beta-1}},
& for $x\in(0,1)$\cr
\cr
0,& otherwise\cr
}$$
where $\alpha>0$ and $\beta>0$.

\beginsection 7a.

Find the mean and variance of $X$.

\bigskip
Solution: This is a beta distribution hence
$$\mu={\alpha\over\alpha+\beta},\qquad
\sigma^2={\alpha\beta\over(\alpha+\beta+1)(\alpha+\beta)^2}$$

\beginsection 7b.

What is the distribution of $X$ whan $\alpha=\beta=1$?

\bigskip
Solution:
$$f(x)={\Gamma(2)\over\Gamma(1)\Gamma(1)}=1$$
$$F(t)=\cases{
0,&$t<0$\cr
\cr
\displaystyle{\int_0^tdt=t},&$0\le t<1$\cr
\cr
1,&$1\le t$\cr
}$$

\beginsection 7c and 7d -- Please see separate file in Digital Dropbox.




\end
