\magnification=1200
\raggedright
\parindent=0pt
Determine whether the set $\{x^2+2x-7,x^2-3,x+1\}$ is linearly independent.

\bigskip
We need to determine if there is a linear combination of the set
elements that sums to zero, like this:
$$c_1(x^2+2x-7)+c_2(x^2-3)+c_3(x+1)=0$$
Multiplying it out and then collecting like terms we have
$$(c_1+c_2)x^2+(2c_1+c_3)x-7c_1-3c_2+c_3=0$$
This is equivalent to solving the following equation:
$$
\left(
\matrix{
1&1&0\cr
2&0&1\cr
-7&-3&1
}
\right)
\left(
\matrix{
c_1\cr c_2\cr c_3}
\right)
=
\left(\matrix{0\cr0\cr0}\right)
$$
Call the square matrix $A$.
The determinant of $A$ by cofactor exapansion of row 1 is
$1(0+3)-1(2+7)=-6$.
The nonzero determinant indicates that the inverse of $A$ exists.
Therefore we can multiply both sides of the equation by $A^{-1}$ to
obtain
$$
\left(
\matrix{
c_1\cr c_2\cr c_3}
\right)
=
\left(\matrix{0\cr0\cr0}\right)
$$
This is a trivial solution so we conclude that there is no linear combination
of the set elements that sums to zero. Therefore the set elements are indeed
linearly
independent.
\vfill\eject
$$A=\left(\matrix{
1&2&1&3\cr
2&4&3&7\cr
3&6&2&8\cr
}\right)
$$
Find a column basis for $A$.
\bigskip
The column vectors in $A$ span a column space.
By row reducing the column vectors we can find the minimum set of vectors that
span the same space.
$$\pmatrix{
1&2&1&3\cr
2&4&3&7\cr
3&6&2&8\cr
}^T
=
\pmatrix{
1&2&3\cr
2&4&6\cr
1&3&2\cr
3&7&8\cr
}
$$
$R_2-2R_1\rightarrow R_2$

$R_3-R_1\rightarrow R_3$

$R_4-3R_1\rightarrow R_4$
$$
\pmatrix{
1&2&3\cr
0&0&0\cr
0&1&-1\cr
0&1&-1\cr
}
$$
$R_4-R_3\rightarrow R_4$
$$
\pmatrix{
1&2&3\cr
0&0&0\cr
0&1&-1\cr
0&0&0\cr
}
$$
Transposing the result, the column basis is the set of nonzero column vectors.
$$
\left\{
\pmatrix{1\cr2\cr3}
,
\pmatrix{0\cr1\cr-1}
\right\}
$$
\vfill\eject
$$A=\left(\matrix{
1&2&1&3\cr
2&4&3&7\cr
3&6&2&8\cr
}\right)
$$
Find a row basis for $A$.
\bigskip
The row vectors in $A$ span a row space.
By row reduction we can find the minimum set of vectors
that span the same space.
\bigskip
$R_2-2R_1\rightarrow R_2$

$R_3-3R_1\rightarrow R_3$
$$\pmatrix{
1&2&1&3\cr
0&0&1&1\cr
0&0&-1&-1\cr
}$$
$R_3+R2\rightarrow R_3$
$$\pmatrix{
1&2&1&3\cr
0&0&1&1\cr
0&0&0&0\cr
}$$
The set of nonzero row vectors is the row basis.
$$\left\{
\pmatrix{1\cr2\cr1\cr3}
,
\pmatrix{0\cr0\cr1\cr1}
\right\}
$$
\vfill\eject
$$A=\left(\matrix{
1&2&1&3\cr
2&4&3&7\cr
3&6&2&8\cr
}\right)
$$
Find a basis for the null space of $A$.
\bigskip
The null space of $A$ is the set of all vectors $x$ such that $Ax=0$.
From the row space calculation, perform one additional elementary
row operation.
$$\pmatrix{
1&2&1&3\cr
0&0&1&1\cr
0&0&0&0\cr
}$$
$R_1-R_2\rightarrow R_1$
$$
\pmatrix{
1&2&0&3\cr
0&0&1&1\cr
0&0&0&0\cr
}$$
This corresponds to the system of equations
$$
\left\{{\eqalign{
x+2y+2w&=0\cr
z+w&=0}}\right.
$$
Let $y=s$ and $w=t$ to obtain the solution set
$\{(-2s-2t,s,-t,t)^\tau:s,t\in\cal R\}$.
Multiply by $A$ to check the result.
$$
\pmatrix{
1&2&1&3\cr
2&4&3&7\cr
3&6&2&8\cr
}
\pmatrix{
-2s-2t\cr
s\cr
-t\cr
t}
=
\pmatrix{
-2s-2t+2s-t+3t\cr
-4s-4t+4s-3t+7t\cr
-6s-6t+6s-2t+8t
}
=
\pmatrix{0\cr0\cr0}
$$

\end
