\parindent=0pt

Abstract Algebra Homework 6

\bigskip
{\bf Problem A.} Prove the following corollary, using the lemma:
There is no integer between 0 and 1.
\medskip
{\bf Corollary.} Let $a\in Z$. Then there is no integer
between $a$ and $a+1$.

\bigskip
{\bf Proof.} Let $a\in Z$.
Suppose there is a $b\in Z$ such that $a<b<a+1$.
Then by Lemma 2.11 (Unit 7) we can add $-a$ to obtain
$0<b-a<1$.
We have $b-a$ is an integer by closure.
We also have $b-a$ is not an integer because it is between
0 and 1.
This is a contradiction
hence there is no integer between $a$ and $a+1$.

\bigskip
The following proof uses Corollary 1.2, the number 1 is the smallest element of $N$.
\medskip
{\bf Proof.} Let $a\in Z$.
Suppose there is a $b\in Z$ such that $a<b$ and $b<a+1$.
By the first less-than we have $b-a\in N$.
Adding $-a$ to the second less-than yields $b-a<1$.
Then by Corollary 1.2
we have $b-a\not\in N$.
This is a contradiction
hence there is no integer between $a$ and $a+1$.

\vfill
\eject

{\bf Problem B.} Prove that
$2+4+\cdots+2n=n(n+1)$
for any $n\in N$.

\bigskip
{\bf Proof.}
(i) For $n=1$ we have $2=1(1+1)$ hence the equation is true
for $n=1$.
(ii)~Suppose the equation holds for a natural number $n$.
Then we can show that the equation holds for $n+1$ by adding
the next term in the sequence and checking the result.
$$n(n+1)+2(n+1)=(n+1)((n+1)+1)$$
$$n^2+n+2n+2=(n+1)(n+2)$$
$$n^2+3n+2=n^2+3n+2$$
Therefore the equation is true for $n+1$.
Hence by induction the equation is true for all $n\in N$.

\vfill
\eject

{\bf Problem C.} Prove that
$$1+3+9+\cdots+3^n={3^{n+1}-1\over 2}$$
for all $n\in N$.

\bigskip
{\bf Proof.} (i) For $n=1$ we have
$$1+3^1={3^2-1\over 2}=4$$
hence the equation is true for $n=1$.
(ii)~Suppose the equation holds for a natural number $n$.
Then we can show that the equation holds for $n+1$ by
adding the next term in the sequence and checking the result.
$$
{3^{n+1}-1\over2}+3^{n+1}
={3^{(n+1)+1}-1\over 2}
$$
$$
{3^{n+1}-1+2(3^{n+1})\over2}
={3^{n+2}-1\over 2}
$$
$$
{3(3^{n+1})-1\over2}
={3^{n+2}-1\over 2}
$$
$$
{3^{n+2}-1\over2}
={3^{n+2}-1\over 2}
$$
Therefore the equation is true for $n+1$.
Hence by induction the equation is true for all $n\in N$.

\end
