\index{built-in functions}

\section*{abs}
abs($x$) returns the absolute value or vector length of $x$.
The mag function should be used for complex $x$.

\medskip
{\tt P=(x,y)}

{\tt abs(P)}

$$(x^2+y^2)^{1/2}$$

\section*{adj}
adj($m$) returns the adjunct of matrix $m$.

\section*{and}
and($a,b,\ldots$) returns the logical ``and'' of predicate expressions.

\section*{arccos}
arccos($x$) returns the inverse cosine of $x$.

\section*{arccosh}
arccosh($x$) returns the inverse hyperbolic cosine of $x$.

\section*{arcsin}
arcsin($x$) returns the inverse sine of $x$.

\section*{arcsinh}
arcsinh($x$) returns the inverse hyperbolic sine of $x$.

\section*{arctan}
arcttan($x$) returns the inverse tangent of $x$.

\section*{arctanh}
arctanh($x$) returns the inverse hyperbolic tangent of $x$.

\section*{arg}
arg($z$) returns the angle of complex $z$.

\section*{ceiling}
ceiling($x$) returns the smallest integer not less than $x$.

\section*{check}
check($x$) In a script, if the predicate $x$ is true then continue, else stop.

\section*{choose}
choose($n,k$) returns $\displaystyle\left({n \atop k}\right)$

\section*{circexp}
circexp($x$) returns expression $x$ with circular functions converted
to exponential forms.
Sometimes this will simplify an expression.

\section*{coeff}
coeff($p,x,n$) returns the coefficient of $x^n$ in polynomial $p$.

\section*{cofactor}
cofactor($m,i,j$) returns of the cofactor of matrix $m$ with respect to row $i$ and column $j$.

\section*{conj}
conj($z$) returns the complex conjugate of $z$.

\section*{contract}
\index{trace}
contract($a,i,j$) returns tensor $a$ summed over indices $i$ and $j$.
If $i$ and $j$ are omitted then indices 1 and 2 are used.
contract($m$) is equivalent to the trace of matrix $m$.

\section*{cos}
cos($x$) returns the cosine of $x$.
%If $x$ is a floating point number then $\cos(x)$ is evaluated numerically.

\section*{cosh}
cosh($x$) returns the hyperbolic cosine of $x$.

\section*{cross}
cross($u,v$) returns the cross product of vectors $u$ and $v$.

\section*{curl}
curl($u$) returns the curl of vector $u$.

\section*{d}
d($f,x$) returns the derivative of $f$ with respect to $x$.

\section*{defint}
defint($f,x,a,b,\ldots$)
returns the definite integral of $f$ with respect to $x$ evaluated from $a$ to $b$.
The argument list can be extended for multiple integrals.
For example, $d(f,x,a,b,y,c,d)$.

\section*{deg}
deg($p,x$) returns the degree of polynomial $p$ in $x$.

\section*{denominator}
denominator($x$) returns the denominator of expression $x$.

\section*{det}
det($m$) returns the determinant of matrix $m$.

\section*{do}
do($a,b,\ldots$) evaluates the argument list from left to right.
Returns the result of the last argument.

\section*{dot}
dot($a,b,\ldots$) returns the dot product of tensors.

\section*{draw}
draw($f,x$) draws the function $f$ with respect to $x$.

\section*{erf}
erf($x$) returns the error function of $x$.

\section*{erfc}
erf($x$) returns the complementary error function of $x$.

\section*{eval}
eval($f,x,n$) returns $f$ evaluated at $x=n$.

\section*{exp}
exp($x$) returns $e^x$.

\section*{expcos}
expcos($x$) returns the cosine of $x$ in exponential form.

\medskip
{\tt expcos(x)}

$${1\over2}\exp(-ix)+{1\over2}\exp(ix)$$

\section*{expsin}
expsin($x$) returns the sine of $x$ in exponential form.

\medskip
{\tt expsin(x)}

$${1\over2}i\exp(-ix)-{1\over2}i\exp(ix)$$

\section*{factor}
factor($n$) factors the integer $n$.

\medskip
{\tt factor(12345)}

$$3\times 5\times 823$$

\medskip
\noindent
factor($p,x$) factors polynomial $p$ in $x$.
The last argument can be omitted for polynomials in $x$.
The argument list can be extended for multivariate polynomials.
For example, factor($p,x,y$) factors $p$ over $x$ and then over $y$.

\medskip
{\tt factor(125*x{\char94}3-1)}

$$(5x-1)(25x^2+5x+1)$$

\section*{factorial}
Example:

\medskip
{\tt 10!}

$$3628800$$

\section*{filter}
filter($f,a,b,\ldots$) returns $f$ with terms involving $a$, $b$, etc. removed.

\medskip
{\tt 1/a+1/b+1/c}

$${1\over a}+{1\over b}+{1\over c}$$

{\tt filter(last,a)}

$${1\over b}+{1\over c}$$

\section*{float}
float($x$) converts $x$ to a floating point value.

\medskip
{\tt sum(n,0,20,(-1/2){\char94}n)}

$$699051\over1048576$$

{\tt float(last)}

$$0.666667$$

\section*{floor}
floor($x$) returns the largest integer not greater than $x$.

\section*{for}
for($i,j,k,a,b,\ldots$) For $i$ equals $j$ through $k$ evaluate $a$, $b$, etc.

\medskip
{\tt x=0}

{\tt y=2}

{\tt for(k,1,9,x=sqrt(2+x),y=2*y/x)}

{\tt float(y)}

$$3.14159$$

\section*{gcd}
gcd($a,b,\ldots$) returns the greatest common divisor.

\section*{hermite}
hermite($x,n$) returns the $n$th Hermite polynomial in $x$.

\section*{hilbert}
hilbert($n$) returns a Hilbert matrix of order $n$.

\section*{imag}
imag($z$) returns the imaginary part of complex $z$.

\section*{inner}
inner($a,b,\ldots$) returns the inner product of tensors.
Same as the dot product.

\section*{integral}
integral($f,x$) returns the integral of $f$ with respect to $x$.

\section*{inv}
inv($m$) returns the inverse of matrix $m$.

\section*{isprime}
isprime($n$) returns 1 if $n$ is prime, zero otherwise.

\medskip
{\tt isprime(2{\char94}53-111)}

$$1$$

\section*{laguerre}
laguerre($x,n,a$) returns the $n$th Laguerre polynomial in $x$.
If $a$ is omitted then $a=0$ is used.

\section*{lcm}
lcm($a,b,\ldots$) returns the least common multiple.

\section*{legendre}
legendre($x,n,m$) returns the $n$th Legendre polynomial in $x$.
If $m$ is omitted then $m=0$ is used.

\section*{log}
log($x$) returns the natural logarithm of $x$.

\section*{mag}
mag($z$) returns the magnitude of complex $z$.

\section*{mod}
mod($a,b$) returns the remainder of $a$ divided by $b$.

\section*{not}
not($x$) negates the result of predicate expression $x$.

\section*{nroots}
nroots($p,x$) returns all of the roots, both real and complex, of
polynomial $p$ in $x$.
The roots are computed numerically.
The coefficients of $p$ can be real or complex.

\section*{numerator}
numerator($x$) returns the numerator of expression $x$.
%\begin{itemize}
%\item[$\scriptstyle1$]{\tt numerator(a/b+b/a)}
%\item[$\scriptstyle2$]\hspace{50pt} $a^2+b^2$
%\end{itemize}

\section*{or}
or($a,b,\ldots$) returns the logical ``or'' of predicate expressions.

\section*{outer}
outer($a,b,\ldots$) returns the outer product of tensors.

\section*{polar}
polar($z$) converts complex $z$ to polar form.

\section*{prime}
prime($n$) returns the $n$th prime number, $1\le n\le10{,}000$.

\section*{print}
print($a,b,\ldots$) evaluates expressions and prints the results..
Useful for printing from inside a ``for'' loop.

\section*{product}
product($i,j,k,f$) returns $\displaystyle\prod_{i=j}^k f$

\section*{quote}
quote($x$) returns expression $x$ unevaluated.

\section*{quotient}
quotient($p,q,x$) returns the quotient of polynomials in $x$.

\section*{rank}
rank($a$) returns the number of indices that tensor $a$ has.
A scalar has no indices so its rank is zero.

\section*{rationalize}
rationalize($x$) puts everything over a common denominator.

\medskip
{\tt rationalize(a/b+b/a)}

$$a^2+b^2\over ab$$

\section*{real}
real($z$) returns the real part of complex $z$.

\section*{rect}
rect($z$) returns complex $z$ in rectangular form.

\section*{roots}
roots($p,x$) returns the values of $x$ such that the polynomial $p(x)=0$.
The polynomial should be factorable over integers.

\section*{simplify}
simplify($x$) returns $x$ in a simpler form.

\section*{sin}
sin($x$) returns the sine of $x$.

\section*{sinh}
sinh($x$) returns the hyperbolic sine of $x$.

\section*{sqrt}
sqrt($x$) returns the square root of $x$.

\section*{stop}
In a script, it does what it says.

\section*{subst}
subst($a,b,c$) substitutes $a$ for $b$ in $c$ and returns the result.

\section*{sum}
sum($i,j,k,f$) returns $\displaystyle\sum_{i=j}^k f$

\section*{tan}
tan($x$) returns the tangent of $x$.

\section*{tanh}
tanh($x$) returns the hyperbolic tangent of $x$.

\section*{taylor}
taylor($f,x,n,a$) returns the Taylor expansion of $f$ of $x$ at $a$.
The argument $n$ is the degree of the expansion.
If $a$ is omitted then $a=0$ is used.

\medskip
{\tt taylor(1/cos(x),x,4)}

$${5\over24}x^4+{1\over2}x^2+1$$

\section*{test}
test($a,b,c,d,\ldots$)
If $a$ is true then $b$ is returned else if $c$ is true then $d$ is returned, etc.
If the number of arguments is odd then the last argument is returned when all else fails.

\section*{transpose}
transpose($a,i,j$) returns the transpose of tensor $a$ with respect to indices $i$ and $j$.
If $i$ and $j$ are omitted then 1 and 2 are used.
Hence a matrix can be transposed with a single argument.

\medskip
{\tt A=((a,b),(c,d))}

{\tt transpose(A)}

$$\left(\matrix{a & c\cr b & d\cr}\right)$$

\section*{unit}
unit($n$) returns an $n\times n$ identity matrix.

\medskip
{\tt unit(2)}

$$\left(\matrix{1&0\cr0&1\cr}\right)$$

\section*{zero}
zero($i,j,\ldots$) returns a null tensor with dimensions $i$, $j$, etc.
Useful for creating a tensor and then setting the component values.

\printindex
\end{document}
